---
title: "Homework 3"
author: "CJ Snyder"
date: "10/10/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Uploading Dataset
```{r, results='hide'}
library(p8105.datasets)
data ("instacart")

instacart
```

## **Dataset Description:**
The instacart dataset includes `r nrow(instacart)` observations, along with `r ncol(instacart)` variables including: `r names(instacart)`. 

Each item within an order is linked together using a particular _order_id_ and _product_id_, and includes information like when the order was made (*order_dow*, and *order_hour_of_day*), as well as where the item is located within the store (*`r names(instacart) [12:15]`*) and whether or not this item is being re-ordered (*reordered*). 

The amount of items ordered from each department is summarized in the below bar chart:
```{r, echo=FALSE}
instacart %>% 
  ggplot(aes(x=department)) + geom_bar(stat="count") +
    labs(x="Department where Item was Located", y="Amount Ordered") +
    scale_y_continuous(labels=scales::comma) +
    theme(axis.text.x = element_text(size=12,angle=90))
```

The largest number of orders occured on Sunday according to the below chart:
```{r, echo=FALSE}
instacart %>% 
  ggplot(aes(x=order_dow)) + geom_bar(stat="count") +
    labs(x="Day of the Week", y="Amount Ordered") +
    scale_y_continuous(labels=scales::comma) +
    scale_x_continuous(
      breaks = c(0,1,2,3,4,5,6),
      labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) +
    theme(axis.text.x = element_text(size=12))
```

```{r, echo=FALSE}
aisle_count = 
  instacart %>% 
  group_by(aisle) %>% 
  count() %>% 
  arrange(desc(n)) 
```

The total number of aisles collected by _Instacart_ is `r summarize(instacart, count=n_distinct(aisle_id))`. The aisle that had the most items ordered from it was **`r aisle_count[1,1]`** with *`r aisle_count[1,2]`* items bought, followed closely by **`r aisle_count[2,1]`** with *`r aisle_count[2,2]`* items bought. 

The aisles that had over 10,000 items bought from them are shown in the below plot (organized from most to least):
```{r, echo=FALSE, eval=FALSE}
aisle_count %>% 
  filter(n>10000) %>% 
  ggplot(aes(x=reorder(aisle,desc(n)),y=n)) + 
    geom_col() +
    labs(x="Aisle where Item was Located", y="Amount Ordered") +
    scale_y_continuous(labels=scales::comma) +
    theme(axis.text.x = element_text(size=11,angle=90))
```

```{r, echo=FALSE}
aisle_count %>% 
  filter(n>10000) %>% 
  ggplot(aes(x=reorder(aisle,desc(n)),y=n)) + 
    geom_point() +
    labs(x="Aisle where Item was Located", y="Amount Ordered") +
    scale_y_continuous(labels=scales::comma) +
    theme(axis.text.x = element_text(size=11,angle=90))
```

The three most popular items bought from aisles labeled as _"baking ingredients"_, _"dog food care"_, and _"packaged vegetables fruits"_ are as follows:
```{r, echo=FALSE}
aisle_table = 
  instacart %>% 
  group_by(aisle, product_id, product_name) %>% 
  summarize(count_prod = n()) %>% 
  filter(aisle %in% c("baking ingredients",
                      "dog food care",
                      "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  top_n(3, count_prod) %>% 
  rename("Aisle" = aisle,
         "Amount Bought" = count_prod,
         "Prodcut ID" = product_id,
         "Product Name" = product_name)

print(aisle_table)
```

The mean hour of the day in which _Pink Lady Apples_ and _Coffee Ice Cream_ are ordered throughout the week are listed in the table below:
```{r, echo=FALSE}
mean_hour_df = 
  instacart %>% 
  group_by(product_name, order_dow) %>%
  mutate(
    mean_hour = mean(order_hour_of_day)) %>% 
  filter(product_name %in% c("Pink Lady Apples",
                             "Coffee Ice Cream")) %>% 
  select(product_name, order_dow, mean_hour) %>% 
  distinct(mean_hour) %>% 
  ungroup(order_dow) %>% 
  mutate(
    order_dow = as.character(order_dow),
    order_dow = factor(order_dow, c(0, 1, 2, 3, 4, 5, 6), 
                       labels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
  )
mean_hour_df = 
  mean_hour_df[order(mean_hour_df$order_dow), ] %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) 
print(mean_hour_df)
```


