---
title: "Homework 3"
author: "CJ Snyder"
date: "10/10/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Uploading Dataset
```{r, results='hide'}
library(p8105.datasets)
data ("instacart")

instacart
```

The instacart dataset includes `r nrow(instacart)` observations, along with `r ncol(instacart)` variables including: `r names(instacart)`. 

Each item within an order is linked together using a particular _order_id_ and _product_id_, and includes information like when the order was made (*order_dow*, and *order_hour_of_day*), as well as where the item is located within the store (*`r names(instacart) [12:15]`*) and whether or not this item is being re-ordered (*reordered*). 

The amount of items ordered from each department is summarized in the below bar chart:
```{r, echo=FALSE}
instacart %>% 
  ggplot(aes(x=department)) + geom_bar(stat="count") +
    labs(x="Department where Item was Located", y="Amount Ordered") +
    scale_y_continuous(labels=scales::comma) +
    theme(axis.text.x = element_text(size=12,angle=90))
```


